Instructions for ACL 2020 Proceedings

### 2019-12-06 reports/ACL2020 Paper/acl2020-paper.tex

Dialogue, especially spoken dialogue, is radically different from the kind of data that neural language models are prestrained on.
Not only is the internal structure of contributions different---with features such as disfluencies, repair, incomplete sentences, and various vocal sounds---but the sequential structure of the discourse is also very different.
In dialogue speakers take turns and switch perspectives.
More generally, the way that utterances contribute and cohere to the discourse is different from the relationship between sentences in a piece of written text.

One important dialogical feature missing from text data is laughter.
In the Switchboard dialogue corpus, laughter appears in approximately \%X of utterances.
Laughter also relates to the discourse structure of dialogue. 
% Laughables.
Also useful for predicting DAs (is that already a result?)

@@

\begin{figure}
  \centering
  \includegraphics[width=\the\columnwidth]{img/box-swda.pdf}
  \caption{Box plot for utterances associated with laughter for SWDA. On the left: proportion of DAs having laughter in one of the adjacent utterances, on the right: proportion of DAs containing laughter. }
    \label{fig:box-swda}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\the\columnwidth]{img/box-ami.pdf}
  \caption{Box plot for utterances associated with laughter for AMI-DA. On the left: proportion of DAs having laughter in one of the adjacent utterances, on the right: proportion of DAs containing laughter. }
    \label{fig:box-ami}
\end{figure}

@@

\section{Experiments}
\subsection{Experiment 1: Impact of laughter} \label{sec:experiment1}   % Vlad
In the first experiment we investigated whether laughter, as an example of a dialogue-specific signal, is a helpful feature for DAR.
Therefore, we train another version of each model: one containing laughs (\texttt{L}) and one with laughs left out (\texttt{NL}), and compare their performances in DAR task.
Table~\ref{table:laughter-total-acc}) compares the results from applying the models with two different utterance encoders (\texttt{BERT}, \texttt{CNN}). From observing the total accuracy scores, models appear to be almost unaffected by leaving out the laughters. On the other hand, we can observe differences in performance depending on the dialogue act, regardless how often laughter occurs in the current or adjacent utterances (see Figure~\ref{fig:swda-by-da} for Switchboard and Figure~\ref{fig:ami-by-da} for AMI corpus).

In order to further investigate the differences, we looked at the dialogue act groups.
\begin{table}
  \centering
  \begin{tabular}{@{}lll@{}}
    \toprule
                      & SWDA  & AMI-DA   \\ \midrule
    \texttt{BERT-NL}  & 77.07 & 67.06       \\ 
    \texttt{BERT-L}   & 76.93 & 67.12       \\ \midrule
    \texttt{CNN-NL}   & 75.08 & 63.46        \\
    \texttt{CNN-L}    & 75.40 & 64.30        \\ \bottomrule

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{img/SWDA-bertLvsNL.pdf}
  \caption{Change in accuracy for each SWDA dialogue act. Positive changes when adding laughter are shown in blue. Vertical bars indicate how often dialogue act is associated with laughter.}
    \label{fig:swda-by-da}
\end{figure*}

\begin{figure*}
  \label{fig:ami-by-da}
  \centering
  \includegraphics[width=\textwidth]{img/AMI-DA-bertLvsNL.pdf}
  \caption{Change in accuracy for each AMI dialogue act. Positive changes when adding laughter are shown in blue. Vertical bars indicate how often dialogue act is associated with laughter.}
\end{figure*}
% TODO per dialogue act analysis
% DONE compare impacts of CNN vs BERT
% DONE for both AMI and SWDA
% ?

@@

%\subsection{Experiment 4: Cross-domain pre-training (if time permits)} \label{sec:experiment4} % Bill
%  - cross-domain pre-training
\section{Related work} % add references first
Dialogue act recoginition
\begin{itemize}
  \item \citet{botheContextbasedApproachDialogue2018} -- uses an RNN for context (recently SOTA on SWDA)
  \item \citet{pragstVectorRepresentationUtterances2018} -- dialogue vector models. Uses vector-manual DA tag correspondance to argue for semantic relevance of their vector models (related to how we use DA tagging to verrify BERT usefulness)
  \item \citet{cerisaraEffectsUsingWord2vec2017} -- standard pre-trained word embeddings do not help for DAR in any of three tested languages (English, French, Czech)
  \item \citet{kalchbrennerRecurrentConvolutionalNeural2013} -- DAR model similar to our architecture. Uses an RNN ``discourse model'' and a CNN similar to our baseline encoder
  \item \citet{khanpourDialogueActClassification2016} -- Deep LSTM utt representation. No notion of context. Finds word vectors are useful.
  \item \citet{tranPreservingDistributionalInformation2017} -- Sequential DA prediction with uncertainty propegation
  \item \citet{chenDialogueActRecognition2017} -- SOTA? SWDA DAR using attention and CRF over utterances
  \item \citet{ortegaNeuralbasedContextRepresentation2017} -- SWDA DAR with attention-based context representations and CNN utterance representations
  \item \citet{shenNeuralAttentionModels2016} -- Uses attention for utt encoding in DAR. No notion of context, I think.
  \item \citet{tranHierarchicalNeuralModel2017} -- RNNs at both the utteranec and dialogue level. Also attention. Very similar to our architecture.
\end{itemize}

>>>>

Dialogue, especially spoken dialogue, is radically different from the kind of data that neural language models are prestrained on.
On the example, from the Switchboard corpus shown in Table~\ref{table:example} it is evident that the syntactic structure of dialogue is very different from the written text, even if one imagines a script for a movie or a play.
Not only is the internal structure of contributions different---with features such as disfluencies, repair, incomplete sentences, and various vocal sounds---but the sequential structure of the discourse is also very different.
In dialogue speakers take turns and switch perspectives.
More generally, the way that utterances contribute and cohere to the discourse is different from the relationship between sentences in a piece of written text.

One important dialogical feature missing from text data is laughter.
In the Switchboard dialogue corpus, laughter comes about every 200 tokens.
Laughter also relates to the discourse structure of dialogue. 
% Laughables.
Also useful for predicting DAs (is that already a result?)

\begin{table}
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
    A	&sd	& Well, I'm the kind of cook that I don't normally measure things,  \\
    A	&sd	& I just kind of throw them in \\
    A	&sd	& and, you know, I don't to the point of, you know, measuring down to the exact amount that they say.  \\
    B	&sv	& That means you're a real cook. \\
    A	&bd	& \texttt{<Laughter>} Oh, is that what it means.  \\
    A	&b	& Uh-huh.  \\
    A	&x	& \texttt{<Laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw2827). Dialogue acts: \emph{sd}---Statement-non-opinion, \emph{sv}---Statement-opinion, \emph{bd}---Downplayer, \emph{b}---Backchannel, \emph{x}---Non-verbal. }
  \label{table:example}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{img/box-swda.pdf}
  \caption{Box plot for utterances associated with laughter in SWDA. On the left: proportion of DAs having laughter in one of the adjacent utterances, on the right: proportion of DAs containing laughter. }
    \label{fig:box-swda}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{img/box-ami.pdf}
  \caption{Box plot for utterances associated with laughter in AMI-DA. On the left: proportion of DAs having laughter in one of the adjacent utterances, on the right: proportion of DAs containing laughter. }
    \label{fig:box-ami}
\end{figure}

@@

\section{Experiments}
\subsection{Experiment 1: Impact of laughter} \label{sec:experiment1}   % Vlad
In the first experiment we investigated whether laughter, as an example of a dialogue-specific signal, is a helpful feature for DAR.
Therefore, we train another version of each model: one containing laughs (\texttt{L}) and one with laughs left out (\texttt{NL}), and compare their performances in DAR task.
Table~\ref{table:laughter-total-acc}) compares the results from applying the models with two different utterance encoders (\texttt{BERT}, \texttt{CNN}).
First of all, we can see that \texttt{BERT} utterance encoder outperforms \texttt{CNN} in all the cases.
From observing the effect of laughters, we can see differences in performance depending on the dialogue act, regardless how often laughter occurs in the current or adjacent utterances (see Figure~\ref{fig:swda-by-da} for Switchboard and Figure~\ref{fig:ami-by-da} for AMI corpus).
The strongest evidence for laughter as a helpful feature was found in SWDA for the \texttt{BERT} utterance encoder, where macro-F1 score increases by 7.89 percentage points.

\begin{table}
  \centering
  \begin{tabular}{@{}lcccc@{}}
    \toprule
                      & \multicolumn{2}{c}{SWDA} & \multicolumn{2}{c}{AMI-DA} \\ \midrule
                      & macro-F1 & acc. & macro-F1 & acc.       \\ 
    \texttt{BERT-NL}  & 38.10 & 77.07 & 49.09 & 67.06       \\ 
    \texttt{BERT-L}   & 45.99 & 76.93 & 50.17 & 67.12       \\ \midrule
    \texttt{CNN-NL}   & 37.23 & 75.08 & 38.37 & 63.46        \\
    \texttt{CNN-L}    & 27.59 & 75.40 & 37.94 & 64.30        \\ \bottomrule

@@

In order to further investigate the differences, we looked at the dialogue act groups.
Mapping all dialogue acts into four groups allows direct comparison between model performances in SWDA and AMI-DA.
Table~\ref{table:laughter-group-acc} compares accuracies between the two main groups: Forward-Communicative-Function and Backward-Communicative-Function (82.3\% of all DAs in SWDA and 76.7\% in AMI-DA).
Our models tend to perform better for Forward-Communicative-Function group irrespective of presence the corpus and the presence of laughs.
Also, our models perform better for AMI-DA corpus.
Overall performance in these groups appears to be unaffected by presence of laughs.

\begin{table}
  \centering
  \begin{tabular}{@{}llcc@{}}
    \toprule
    &                  & Forward-CF  & Backwards-CF   \\ \midrule
    SWDA &\texttt{BERT-NL}  & 92.54 & 91.67       \\ 
         &\texttt{BERT-L}   & 95.42 & 90.91       \\ 
         &\texttt{CNN-NL}   & 93.30 & 90.85        \\
         &\texttt{CNN-L}    & 94.64 & 90.13        \\ \midrule
    AMI  &\texttt{BERT-NL}  & 87.77 & 80.83       \\ 
         &\texttt{BERT-L}   & 87.36 & 79.87       \\
         &\texttt{CNN-NL}   & 91.85 & 71.21        \\
         &\texttt{CNN-L}    & 91.25 & 71.28        \\    \bottomrule
  \end{tabular}
  \caption{Comparison of models depending on using laughter on the training phase. }
  \label{table:laughter-group-acc}
\end{table}

% to appendix?
\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{img/SWDA-bertLvsNL.pdf}
  \caption{Change in accuracy for each SWDA dialogue act. Positive changes when adding laughter are shown in blue. Vertical bars indicate how often dialogue act is associated with laughter.}
    \label{fig:swda-by-da}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{img/AMI-DA-bertLvsNL.pdf}
  \caption{Change in accuracy for each AMI dialogue act. Positive changes when adding laughter are shown in blue. Vertical bars indicate how often dialogue act is associated with laughter.}
    \label{fig:ami-by-da}
  \end{figure*}

%
% TODO per dialogue act analysis
% DONE compare impacts of CNN vs BERT
% DONE for both AMI and SWDA
% ?

@@

%\subsection{Experiment 4: Cross-domain pre-training (if time permits)} \label{sec:experiment4} % Bill
%  - cross-domain pre-training
\section{Related work} % add references first
For the background:

Dialogue act recoginition
\begin{itemize}
  \item \citet{botheContextbasedApproachDialogue2018} -- uses an RNN for context (recently SOTA on SWDA)
  \item \citet{pragstVectorRepresentationUtterances2018} -- dialogue vector models. Uses vector-manual DA tag correspondance to argue for semantic relevance of their vector models (related to how we use DA tagging to verrify BERT usefulness)
  \item \citet{cerisaraEffectsUsingWord2vec2017} -- standard pre-trained word embeddings do not help for DAR in any of three tested languages (English, French, Czech)
  \item \citet{kalchbrennerRecurrentConvolutionalNeural2013} -- DAR model similar to our architecture. Uses an RNN ``discourse model'' and a CNN similar to our baseline encoder
  \item \citet{khanpourDialogueActClassification2016} -- Deep LSTM utt representation. No notion of context. Finds word vectors are useful.
  \item \citet{tranPreservingDistributionalInformation2017} -- Sequential DA prediction with uncertainty propegation
  \item \citet{chenDialogueActRecognition2017} -- SOTA? SWDA DAR using attention and CRF over utterances
  \item \citet{ortegaNeuralbasedContextRepresentation2017} -- SWDA DAR with attention-based context representations and CNN utterance representations
  \item \citet{shenNeuralAttentionModels2016} -- Uses attention for utt encoding in DAR. No notion of context, I think.
  \item \citet{tranHierarchicalNeuralModel2017} -- RNNs at both the utteranec and dialogue level. Also attention. Very similar to our architecture.
\end{itemize}
